---
title: "Template"
author: "Anthony O'Brien"
date: "September 30, 2018"
output: html_document
---

#Data import

```{r}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("RPostgreSQL")
#install.packages("Hmisc")
#install.packages("dplyr")
#install.packages("caret")
#install.packages("sjPlot")
#install.packages("DBI")
#install.packages("viridis")
#install.packages("car")
#install.packages("ggpubr")
library("ggpubr")
library("car")
library("tidyr")
library("gridExtra")
library("grid")
library("viridis")
library("dplyr")
library("ggplot2")
library("bigrquery")
library("DBI")
library("RPostgreSQL")
library("devtools")
library("MIMICbook")
```

```{r, include = FALSE}
if(!("devtools" %in% installed.packages()[,1])) {install.packages("devtools",repos="https://cloud.r-project.org")}
if(!("MIMICbook" %in% installed.packages()[,1])) {install_github("jraffa/MIMICbook")}
```

```{r dbconnect, include=FALSE}
# Load configuration settings
project_id <- "hst-953-2018"
options(httr_oauth_cache=TRUE)
run_query <- function(query)
    {
    data <- query_exec(query, project=project_id, max_pages = Inf, use_legacy_sql = FALSE)
    return(data)
    }
```

```{r load_data, include=TRUE}
#sql query for mimic
sql_query_mimic <- "SELECT
sub.*,
sofa.SOFA,
d.ICD9_CODE,
dicd.LONG_TITLE,
dicd.SHORT_TITLE
FROM (WITH ce as (
 select 
 ce.hadm_id, 
 (case when itemid in (211,220045) and valuenum > 0 and valuenum < 300 then valuenum else null end) as HeartRate
 FROM `physionet-data.mimiciii_clinical.chartevents` ce
  -- exclude rows marked as error
 where (ce.error IS NULL OR ce.error = 1)
 and ce.itemid in (211, 220045) ##Heart Rate
 ),
 
pvt as (
      select
      ce.HeartRate,
      le.hadm_id,
      case when itemid = 50800 then 'SPECIMEN'
         #when itemid = 50801 then 'AADO2'
         when itemid = 50802 then 'BASEEXCESS'
         #when itemid = 50803 then 'BICARBONATE'
         #when itemid = 50804 then 'TOTALCO2'
         #when itemid = 50805 then 'CARBOXYHEMOGLOBIN'
         #when itemid = 50806 then 'CHLORIDE'
         #when itemid = 50808 then 'CALCIUM'
         #when itemid = 50809 then 'GLUCOSE'
         #when itemid = 50810 then 'HEMATOCRIT'
         #when itemid = 50811 then 'HEMOGLOBIN'
         #when itemid = 50812 then 'INTUBATED'
         #when itemid = 50813 then 'LACTATE'
         #when itemid = 50814 then 'METHEMOGLOBIN'
         #when itemid = 50815 then 'O2FLOW'
         when itemid = 50816 then 'FIO2'
         when itemid = 50817 then 'SO2' 
         when itemid = 50818 then 'PCO2'
         when itemid = 50819 then 'PEEP'
         when itemid = 50820 then 'PH'
         when itemid = 50821 then 'PO2'
         #when itemid = 50822 then 'POTASSIUM'
         #when itemid = 50823 then 'REQUIREDO2'
         #when itemid = 50824 then 'SODIUM'
         when itemid = 50825 then 'TEMPERATURE'
         #when itemid = 50826 then 'TIDALVOLUME'
         #when itemid = 50827 then 'VENTILATIONRATE'
         #when itemid = 50828 then 'VENTILATOR'
         else null end as label,
      charttime, 
      value, 
         case when valuenum <= 0 then null
        #when itemid = 50810 and valuenum > 100 then null # Hematocrit
         when itemid = 50816 and valuenum < 20 then null # FIO2
         when itemid = 50816 and valuenum > 100 then null # FIO2
         when itemid = 50817 and valuenum > 100 then null # O2 sat
        #when itemid = 50815 and valuenum > 70 then null # O2 flow
         when itemid = 50821 and valuenum > 800 then null # PO2
         else valuenum
         end as valuenum
         from `physionet-data.mimiciii_clinical.labevents` le
         left join ce
         on le.hadm_id = ce.hadm_id
         where le.ITEMID in (
    50800, 50801, 50802, 50803, 50804, 50805, 50806, 50807, 50808, 50809
   , 50810, 50811, 50812, 50813, 50814, 50815, 50816, 50817, 50818, 50819
   , 50820, 50821, 50822, 50823, 50824, 50825, 50826, 50827, 50828
   , 51545)),
   
   grp as (
     select 
     pvt.hadm_id, 
     pvt.charttime, 
     avg(pvt.HeartRate) as HeartRate,
     max(case when label = 'SPECIMEN' then value else null end) as SPECIMEN,
     #avg(case when label = 'AADO2' then valuenum else null end) as AADO2, 
     avg(case when label = 'BASEEXCESS' then valuenum else null end) as BASEEXCESS, 
     #avg(case when label = 'BICARBONATE' then valuenum else null end) as BICARBONATE, 
     #avg(case when label = 'TOTALCO2' then valuenum else null end) as TOTALCO2,
     #avg(case when label = 'CARBOXYHEMOGLOBIN' then valuenum else null end) as CARBOXYHEMOGLOBIN,
     #avg(case when label = 'CHLORIDE' then valuenum else null end) as CHLORIDE,
     #avg(case when label = 'CALCIUM' then valuenum else null end) as CALCIUM,
     #avg(case when label = 'GLUCOSE' then valuenum else null end) as GLUCOSE,
     #avg(case when label = 'HEMATOCRIT' then valuenum else null end) as HEMATOCRIT,
     #avg(case when label = 'HEMOGLOBIN' then valuenum else null end) as HEMOGLOBIN,
     #avg(case when label = 'INTUBATED' then valuenum else null end) as INTUBATED,
     #avg(case when label = 'LACTATE' then valuenum else null end) as LACTATE,
     #avg(case when label = 'METHEMOGLOBIN' then valuenum else null end) as METHEMOGLOBIN,
     #avg(case when label = 'O2FLOW' then valuenum else null end) as O2FLOW,
     avg(case when label = 'FIO2' then valuenum else null end) as FIO2,
     avg(case when label = 'SO2' then valuenum else null end) as SPO2,
     avg(case when label = 'PCO2' then valuenum else null end) as PACO2, 
     avg(case when label = 'PEEP' then valuenum else null end) as PEEP,
     avg(case when label = 'PH' then valuenum else null end) as PH,
     avg(case when label = 'PO2' then valuenum else null end) as PAO2,
     #avg(case when label = 'POTASSIUM' then valuenum else null end) as POTASSIUM,
     #avg(case when label = 'REQUIREDO2' then valuenum else null end) as REQUIREDO2,
     #avg(case when label = 'SODIUM' then valuenum else null end) as SODIUM,
     avg(case when label = 'TEMPERATURE' then valuenum else null end) as TEMPERATURE
     #avg(case when label = 'TIDALVOLUME' then valuenum else null end) as TIDALVOLUME,
     #max(case when label = 'VENTILATIONRATE' then valuenum else null end) as VENTILATIONRATE,
     #max(case when label = 'VENTILATOR' then valuenum else null end) as VENTILATOR
     from pvt 
     group by pvt.hadm_id, pvt.charttime
     having sum(case when label = 'SPECIMEN' then 1 else 0 end)<2)

select 
adm.subject_id,
icu.icustay_id,
adm.ETHNICITY,
pats.GENDER,
DATETIME_DIFF(adm.admittime, pats.dob, YEAR) AS age,
CASE WHEN DATETIME_DIFF(adm.admittime, pats.dob, YEAR) <=18 then 1 ELSE 0 END AS exclusion_age, # age exclusion
#CASE WHEN DATETIME_DIFF(adm.admittime, pats.dob, YEAR) >120 then 1 ELSE 0 END AS ninety_or_older,
RANK() OVER (PARTITION BY icu.subject_id ORDER BY icu.intime) AS icustay_id_order,
CASE WHEN (RANK() OVER (PARTITION BY icu.subject_id ORDER BY icu.intime)) >1 then 1 ELSE 0 END AS exclusion_stay, #only 1 unique visit or exclude
DATETIME_DIFF(icu.outtime, icu.intime, DAY) AS icu_length_of_stay_day,
CASE WHEN DATETIME_DIFF(icu.outtime, icu.intime, DAY) <(4) OR NULL then 1 ELSE 0 END AS exclusion_length_of_stay, #less than 4 hours exclusion
CASE WHEN grp.HeartRate IS NULL THEN 1 ELSE 0 END AS exclusion_hr, #exclusion hr criteria
pats.dob,
grp.*
from grp
inner join `physionet-data.mimiciii_clinical.admissions` adm
on grp.hadm_id = adm.hadm_id
left join `physionet-data.mimiciii_clinical.icustays` icu
on grp.hadm_id = icu.hadm_id
left join `physionet-data.mimiciii_clinical.patients` pats
on adm.subject_id = pats.subject_id
order by grp.hadm_id, grp.charttime) sub


#INNER JOIN `physionet-data.mimiciii_derived.oasis` oasis
#ON sub.hadm_id=oasis.hadm_id
JOIN `physionet-data.mimiciii_clinical.diagnoses_icd` d
ON sub.hadm_id=d.hadm_id
JOIN `physionet-data.mimiciii_clinical.d_icd_diagnoses` dicd
ON d.ICD9_CODE=dicd.ICD9_CODE
JOIN `physionet-data.mimiciii_derived.sofa` sofa
ON sub.hadm_id=sofa.hadm_id


WHERE exclusion_age = 0
AND exclusion_stay = 0
AND exclusion_length_of_stay != 1
AND exclusion_hr = 0
AND sub.icu_length_of_stay_day IS NOT NULL"
 
```

Please note to optimize speed the above code already applies the exclusion 
criterias, however if you want to run from the original raw data
you can remove the last 5 lines of code in the SQL query. This is why we 
look at the number of users in the dataframe data1 and then mimic_bg.

```{r}
#place mimic into a dataframe
data1 <- run_query(sql_query_mimic)
```

```{r}
#number of unique subjects and their entries
data1 %>% distinct(subject_id, .keep_all = TRUE) %>% tally()
```

# Data preparation and testing (cleaning)

```{r}
#I group the data by hadm_id, then tally/count
#then this information is stored into a dataframe.
#to understand the distribution of the users later on.
n_admissions <- tally(group_by(data1, hadm_id)) #idem
```

```{r}
# Clean data1
#In this part I select the rows I want (the inclusion critera), then I drop columns I do not want, then I 
#add on the tally of the ids, which is useful for counting the number of bloog gas the subject has

mimic_bg <-data1 %>% filter(exclusion_age == 0 & exclusion_stay == 0 & exclusion_length_of_stay == 0 & exclusion_hr == 0) %>%
  select(-c(icustay_id, age, exclusion_age, icustay_id_order, exclusion_stay, icu_length_of_stay_day,
              exclusion_length_of_stay, exclusion_hr, dob, charttime, HeartRate, PEEP, FIO2, SPECIMEN)) %>% 
  left_join(y=n_admissions, by=c("hadm_id"))
```

```{r}
#number of unique subjects and their entries
mimic_bg %>% distinct(subject_id, keep.all=TRUE) %>% tally()
```

```{r}
#Test raw curve using available SO2 and PO2 FROM MIMIC
mimic_bg %>% ggplot(aes(PAO2,SPO2)) + 
  geom_jitter(color="orange", shape="8") +
  geom_smooth(color="blue", se=FALSE, method=loess, span=0.5) +
  theme_classic()
``` 

```{r}
#Rename df columns to fit ODC equations
df <- mimic_bg %>% filter(!is.na(PAO2)) %>% rename(PO2=PAO2, TEMP=TEMPERATURE, PCO2=PACO2, BE=BASEEXCESS)

#check number of patients
df %>% distinct(subject_id, keep.all=TRUE) %>% tally()
```

```{r}
# Oxygen disociation standard curve based off Kellman's equations from Adair

po2seq <-0:300

kellman <- function(po2seq)
  {
  (
    100*((-8.532289*10^3*po2seq)+(2.1214010*10^3*po2seq^2)+(-6.7073989*10^1*po2seq^3)+
     (po2seq^4))/((9.3596087*10^5)+(-3.1346258*10^4*po2seq)+
                   (2.3961674*10^3*po2seq^2)+(-6.7104406*10^1*po2seq^3)+(po2seq^4)))
  }

standardso2 <- mapply(kellman, po2seq)

standard_curve <-data.frame(po2seq, standardso2)

```

```{r}
#Test ODC standard curve
standard_curve %>% ggplot(aes(po2seq,standardso2))+
  geom_line(color="red")+
  theme_classic() +
  labs(x="Oxygen partial pressure", 
       y="Oxygen saturation")

```

```{r}
# RUN ONLY ONE TIME
# Add Adair's theoretical curve to the data frame

df["theoretical_so2"]<-sapply(df$PO2, kellman)
```



```{r}
#RUN ONLY ONE TIME
################################################################################
################################################################################
## Kelman's equation (correction for temperature, ph, CO2 using Adair's eq)   ##
## Reference:                                                                 ##
## Kelman, G.R., 1966. Digital computer subroutine for the conversion of oxygen#
## tension into saturation. Journal of Applied Physiology, 21(4), pp.1375-1376.#
################################################################################
###############################################################################t
## Kelman's equation############################################################

#K_PO2v <- x*10^((0.024(37-TEMP))+(0.40*(PH-7.40))+(0.06(log(40)-log(PCO2)))) 
#K_SO2 <- ((-8.532289*10^3*x)+(2.1214010*10^3*x^2)+(-6.7073989*10^1*x^3)+(x^4))/
#  ((9.3596087*10^5)+(-3.1346258*10^4*x)+(2.3961674*10^3*x^2)+
#     (-6.7104406*10^1*x^3)+(x^4))

################################################################################
################################################################################
################################################################################

## STEP 1: calculate the virtual PO2

K_PO2V <- function(PO2,TEMP,PH,PCO2)
  {
  (PO2*(10^((0.024*(37-TEMP))+(0.40*(PH-7.40))+(0.06*(log(40)-log(PCO2)))))) 
  }

KPO2V <- mapply(K_PO2V, df$PO2, df$TEMP, df$PH, df$PCO2)

df<-df %>% mutate(KPO2V = KPO2V)

## STEP 2: Use virtual PO2 in Adair's equation to adjust for temp, pH AND PCO2

K_SO2 <- function(KPO2V)
  {
  (
    100*((-8.532289*10^3*KPO2V)+(2.1214010*10^3*KPO2V^2)+(-6.7073989*10^1*KPO2V^3)+
     (KPO2V^4))/((9.3596087*10^5)+(-3.1346258*10^4*KPO2V)+
                   (2.3961674*10^3*KPO2V^2)+(-6.7104406*10^1*KPO2V^3)+(KPO2V^4)))
  }

KSO2 <- mapply(K_SO2, df$KPO2V)

df<- df %>% mutate(KSO2 = KSO2)

```

```{r}
#RUN ONLY ONE TIME
################################################################################
################################################################################
## Severinghaus' equation                                                     ##
## Reference:                                                                 ##
## Severinghaus, J.W., 1979. Simple, accurate equations for human blood O2    ##
## dissociation computations. Journal of Applied Physiology, 46(3), pp.599-602##
################################################################################
###############################################################################o
## Severinghaus' equation#######################################################

#((((((PO2^3)+(150*PO2))^-1)*23400)+1))^-1)

################################################################################
################################################################################
################################################################################

S_S02 <- function(PO2)
  {
  (((((((PO2^3)+(150*PO2))^-1)*23400)+1)^-1))
  }

SSO2 <- mapply(S_S02, df$PO2)
  
df<- df %>% mutate(SSO2 = SSO2*100)

```

```{r}
#RUN ONLY ONE TIME
################################################################################
################################################################################
## Severinghaus' equation (correction for pH)                                 ##
## Reference:                                                                 ##
## Severinghaus, J.W., 1979. Simple, accurate equations for human blood O2    ##
## dissociation computations. Journal of Applied Physiology, 46(3), pp.599-602##  
################################################################################
###############################################################################b
## Severinghaus' equation (correction for pH)################################### 
#x*((exp(1.1(PH-7.4))))
#LN_S_PO2V_PH <- ((S_PO2V/26.7)^0.184)+(0.003*BE)-(2.2)
#S_PO2V_PH <- exp(LN_S_PO2V_PH)

################################################################################
################################################################################
################################################################################

## STEP 1: Calculate the virtual PO2 corrected for ph according to Severinghaus

S_PO2V <- function(PO2,PH) 
  {
  (PO2*(exp(1.1*(PH-7.4))))
  }

SPO2V <- mapply(S_PO2V, df$PO2, df$PH)

df <-df %>% mutate(SPO2V = SPO2V)

## STEP 2: Use SPO2V and the following equation to account for Bohr effect

LN_SPO2V_PH <- function(SPO2V, BE) 
  {
 ((SPO2V/26.7)^0.184)+(0.003*BE)-(2.2) 
  }

LNSPO2VPH <- mapply(LN_SPO2V_PH, df$SPO2V, df$BE)

df <-df %>% mutate(LNSPO2VPH = LNSPO2VPH*-100)

# S_PO2V_PH <- function(LNSPO2VPH)
#   {
#   exp(LNSPO2VPH)
#   }
# 
# SPO2VPH <- mapply(S_PO2V_PH, df$LNSPO2VPH)
# 
# df <-df %>% mutate(SPO2VPH = SPO2VPH)

## STEP 3: Calculate the SO2 using the corrected PO2 for PH according 
##         to Severinghaus

S_S02_PH <- function(LNSPO2VPH)
  {
  ((((((LNSPO2VPH^3)+(150*LNSPO2VPH))^-1)*23400)+1))^-1
  }

SS02PH <- mapply(S_S02_PH, df$LNSPO2VPH)

df <-df %>% mutate(SS02PH = SS02PH*100)

```

```{r}
#RUN ONLY ONE TIME
################################################################################
################################################################################
## Severinghaus' equation (correction for T)                                  ##
## Reference:                                                                 ##
## Severinghaus, J.W., 1979. Simple, accurate equations for human blood O2    ##
## dissociation computations. Journal of Applied Physiology, 46(3), pp.599-602## 
################################################################################
###############################################################################a
## Severinghaus' equation (correction for T)####################################

#ln_S_PO2V <- (0.058(((0.243*(x/100))^3.88) + 1)^-1)+0.013
#S_PO2V_T <- exp(ln_S_PO2V)
#S_S02_T <- ((((((S_PO2V_T^3)+(150*S_PO2V_T))^-1)*23400)+1))^-1

################################################################################
################################################################################
################################################################################

## Step 1: calculate the virtual PO2 corrected for T according to Severinghaus

ln_S_PO2V <- function(PO2)
  {
  (0.058*(((0.243*(PO2/100))^3.88) + 1)^-1)+0.013
  }

lnSPO2V <- mapply(ln_S_PO2V, df$PO2)

df <-df %>% mutate(lnSPO2V = lnSPO2V*1000)

# S_PO2V_T <- function(lnSPO2V)
#   {
#   exp(lnSPO2V)
#   }
# 
# SPO2VT <- mapply(S_PO2V_T, df$lnSPO2V)

# df <-df %>% mutate(SPO2VT = SPO2VT)

## Step 2: calculate the SO2 using the virtual PO2 according to Severinghaus

S_SO2_T <- function(lnSPO2V)
  {
  ((((((lnSPO2V^3)+(150*lnSPO2V))^-1)*23400)+1))^-1
  }

SSO2T <- mapply(S_SO2_T, df$lnSPO2V)

df <-df %>% mutate(SSO2T = SSO2T*100)

```

#Putting it all together

```{r}

# ODC curves for each model

p1<-ggplot()+
  geom_smooth(data=df, aes(PO2,SPO2), color="#0D1687", se=FALSE)+
  geom_line(data=df, aes(PO2,theoretical_so2), color="#FF332F")+
  labs(x="", 
       y="Oxygen saturation (%)",
       title="", 
       subtitle="")+
  theme_classic()+
  theme(legend.position="none")+
  coord_cartesian(xlim=(c(0,300)))

p2<-ggplot()+
  geom_smooth(data=df, aes(PO2,SPO2), color="#0D1687", se=FALSE, span=0.3)+
  geom_line(data=df, aes(KPO2V,KSO2, color="#FF332F"))+
  labs(x="", 
       y="Oxygen saturation (%)",
       title="", 
       subtitle="")+
  theme_classic()+
  theme(legend.position="none")+
  coord_cartesian(xlim=(c(0,300)))

p3<-ggplot()+
  geom_smooth(data=df, aes(PO2,SPO2), color="#0D1687", se=FALSE, span=0.3)+
  geom_line(data=df, aes(PO2, SSO2, color="#FF332F"))+
  labs(x="", 
       y="O2 Sat (%)",
       title="", 
       subtitle="")+
  theme_classic()+
  theme(legend.position="none")+
  coord_cartesian(xlim=(c(0,300)))

p4<-ggplot()+
  geom_smooth(data=df, aes(PO2,SPO2), color="#0D1687", se=FALSE, span=1)+
  geom_line(data=df, aes(LNSPO2VPH, SS02PH, color="#FF332F"))+
  labs(x="", 
       y="O2 Sat (%)",
       title="", 
       subtitle="")+
  theme_classic()+
  theme(legend.position="none")+
  coord_cartesian(xlim=(c(0,300)))

p5<-ggplot()+
  geom_smooth(data=df, aes(PO2,SPO2), color="#0D1687", se=FALSE, span=1)+
  geom_line(data=df, aes(lnSPO2V, SSO2T, color="#FF332F"))+
  labs(x="Oxygen partial pressure (mmHg)", 
       y="O2 Sat (%)",
       title="", 
       subtitle="")+
  theme_classic()+
  theme(legend.position="none")+
  coord_cartesian(xlim=(c(0,300)))

# Extrapolate the smooth data for patient data, this is bases for comparison of patient data

p<-qplot(df$PO2,df$SPO2) + stat_smooth()
linear_1<-ggplot_build(p)$data[[2]]
linear_1["theoretical_so2"]<-sapply(linear_1$x, kellman)
linear_1["theoretical_so2_x"]<-linear_1$theoretical_so2

# Cumulative distribtion curve

pcd1<-ggplot() + 
  stat_ecdf(data=linear_1, aes(theoretical_so2), geom = "step", color="#FF332F") +  
  stat_ecdf(data=linear_1, aes(y), geom = "step", color="#0D1687") + 
  labs(x="", 
       y="Cumulative density",
       title="")+ 
  theme_classic()

pcd2<-ggplot() + 
  stat_ecdf(data=df, aes(KSO2), geom = "step", color="#FF332F") +
  stat_ecdf(data=linear_1, aes(y), geom = "step", color="#0D1687")  +
  labs(x="", 
       y="Cumulative density",
       title="")+ 
  theme_classic()

pcd3<-ggplot() + 
  stat_ecdf(data=df, aes(SSO2), geom = "step", color="#FF332F") +
  stat_ecdf(data=linear_1, aes(y), geom = "step", color="#0D1687") +
  labs(x="", 
       y="CD",
       title="")+ 
  theme_classic()

pcd4<-ggplot() + 
  stat_ecdf(data=df, aes(SS02PH), geom = "step", color="#FF332F") +
  stat_ecdf(data=linear_1, aes(y), geom = "step", color="#0D1687") +
  labs(x="", 
       y="CD",
       title="")+ 
  theme_classic() 

pcd5<-ggplot() + 
  stat_ecdf(data=df, aes(SSO2T), geom = "step", color="#FF332F") +
  stat_ecdf(data=linear_1, aes(y), geom = "step", color="#0D1687") +
  labs(x="Oxygen saturation (%)", 
       y="CD",
       title="")+ 
  theme_classic()  


# Kolmogorov Smirnov test

ks.test(linear_1$y, df$theoretical_so2)
ks.test(linear_1$y,df$KSO2)
ks.test(linear_1$y,df$SSO2)
ks.test(linear_1$y,df$SS02PH)
ks.test(linear_1$y,df$SSO2T)

# Define number of samples and permutations

N<-10000
B <- 10000

# Define sampling function for each variable

f1 <- function(y){
  X <- sample(df$theoretical_so2, N, replace = TRUE)
  mean(X)
}

f1b<-function(y){
  X <- sample(linear_1$theoretical_so2, N, replace = TRUE)
  mean(X)
}

f2 <- function(y){
  X <- sample(linear_1$y, N, replace = TRUE)
  mean(X)
}

df3<- df %>% filter(!is.na(KSO2)) %>% select(KSO2)
f3 <- function(y){
  X <- sample(df3$KSO2, N, replace = TRUE)
  mean(X)
}

df4<- df %>% filter(!is.na(SSO2)) %>% select(SSO2)
f4 <- function(y){
  X <- sample(df4$SSO2, N, replace = TRUE)
  mean(X)
}

df5<- df %>% filter(!is.na(SS02PH)) %>% select(SS02PH)
f5 <- function(y){
  X <- sample(df5$SS02PH, N, replace = TRUE)
  mean(X)
}

df6<-df %>% filter(!is.na(SSO2T)) %>% select(SSO2T)
f6 <- function(y){
  X <- sample(df6$SSO2T, N, replace = TRUE)
  mean(X)
}

# Monte Carlo permutation for modeled data and theoeretical data 

R1 <- as.data.frame(replicate(B, f1(x)))
R1b <- as.data.frame(replicate(B, f1b(x)))
R2 <- as.data.frame(replicate(B, f2(x)))
R3 <- as.data.frame(replicate(B, f3(x)))
R4 <- as.data.frame(replicate(B, f4(x)))
R5 <- as.data.frame(replicate(B, f5(x)))
R6 <- as.data.frame(replicate(B, f5(x)))

# Rename columns 
names(R1)[1]<-"theoretical"
names(R1b)[1]<-"theoretical_for_patient"
names(R2)[1]<-"patient"
names(R3)[1]<-"kellman"
names(R4)[1]<-"severinghaus_unadjusted"
names(R5)[1]<-"severinghaus_ph"
names(R6)[1]<-"severinghaus_temp"


# Histograms 
# note the histograms need to be adjusted for comparing patient data 
# to models. At present the individual plots are looking at the models 
# comparing each other while the final one compares to patients.

ph1<-ggplot()+
  geom_histogram(aes(R1b$theoretical_for_patient), fill="#FF332F", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R2$patient), fill="#0D1687", color="grey10", alpha=1/10, bins = 200)+
  labs(x="Oxygen saturation (%)", 
       y="Count",
       title="")+ 
  theme_classic()

ph2<-ggplot()+
  geom_histogram(aes(R1$theoretical), fill="#FF332F", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R3$kellman), fill="#0D1687", color="grey10", alpha=1/10, bins = 200)+
  labs(x="Oxygen saturation (%)", 
       y="Count",
       title="")+ 
  theme_classic()

ph3<-ggplot()+
  geom_histogram(aes(R1$theoretical), fill="#FF332F", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R4$severinghaus_unadjusted), fill="#0D1687", color="grey10", alpha=1/10, bins = 200)+
  labs(x="Oxygen saturation (%)", 
       y="Count",
       title="")+ 
  theme_classic()

ph4<-ggplot()+
  geom_histogram(aes(R1$theoretical), fill="#FF332F", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R5$severinghaus_ph), fill="#0D1687", color="grey10", alpha=1/10, bins = 200)+
  labs(x="Oxygen saturation (%)", 
       y="Count",
       title="")+ 
  theme_classic()

ph5<-ggplot()+
  geom_histogram(aes(R1$theoretical), fill="#FF332F", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R6$severinghaus_temp), fill="#0D1687", color="grey10", alpha=1/10, bins = 200)+
  labs(x="Oxygen saturation (%)", 
       y="Count",
       title="")+ 
  theme_classic()

ph6<-ggplot()+
  geom_histogram(aes(R1$theoretical), fill="#FF332F", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R2$patient), fill="#0D1687", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R3$kellman), fill="#0D1687", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R4$severinghaus_unadjusted), fill="#0D1687", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R5$severinghaus_ph), fill="#0D1687", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R6$severinghaus_temp), fill="#0D1687", color="grey10", alpha=1/10, bins = 200)+
  labs(x="Oxygen saturation (%)", 
       y="Count",
       title="")+ 
  theme_classic()

# ttest
t.test(R2$patient, R1b$theoretical_for_patient)
t.test(R2$patient, R3$kellman)
t.test(R2$patient, R4$severinghaus_unadjusted)
t.test(R2$patient, R5$severinghaus_ph)
t.test(R2$patient, R6$severinghaus_temp)

#95%CI
t.test(R1$theoretical, conf.level = 0.95)
t.test(R2$patient, conf.level = 0.95)
t.test(R3$kellman, conf.level = 0.95)
t.test(R4$severinghaus_unadjusted, conf.level = 0.95)
t.test(R5$severinghaus_ph, conf.level = 0.95)
t.test(R6$severinghaus_temp, conf.level = 0.95)

#Median
median(R1$theoretical)
median(R2$patient)
median(R3$kellman)
median(R4$severinghaus_unadjusted)
median(R5$severinghaus_ph)
median(R6$severinghaus_temp)

#IQR
IQR(R1$theoretical)
IQR(R2$patient)
IQR(R3$kellman)
IQR(R4$severinghaus_unadjusted)
IQR(R5$severinghaus_ph)
IQR(R6$severinghaus_temp)



#this extracts the smoothed saturation results for kellman adjusted data and adds the theortical saturation values
r<-qplot(df$KPO2V,df$KSO2) + stat_smooth()
linear_2<-ggplot_build(r)$data[[2]]
linear_2["theoretical_so2"]<-sapply(linear_2$x, kellman)
linear_2["theoretical_so2_x"]<-linear_2$theoretical_so2

#this extracts the smoothed saturation results for severinghaus uadjusted data and adds the theortical saturation values
a<-qplot(df$PO2,df$SSO2) + stat_smooth()
linear_3<-ggplot_build(a)$data[[2]]
linear_3["theoretical_so2"]<-sapply(linear_3$x, kellman)
linear_3["theoretical_so2_x"]<-linear_3$theoretical_so2

#this extracts the smoothed saturation results for severinghaus adjusted ph data and adds the theortical saturation values
o<-qplot(df$LNSPO2VPH,df$SS02PH) + stat_smooth()
linear_4<-ggplot_build(o)$data[[2]]
linear_4["theoretical_so2"]<-sapply(linear_4$x, kellman)
linear_4["theoretical_so2_x"]<-linear_4$theoretical_so2

#this extracts the smoothed saturation results for severinghaus adjusted t data and adds the theortical saturation values
b<-qplot(df$lnSPO2V,df$SSO2T) + stat_smooth()
linear_5<-ggplot_build(p)$data[[2]]
linear_5["theoretical_so2"]<-sapply(linear_5$x, kellman)
linear_5["theoretical_so2_x"]<-linear_5$theoretical_so2

#Linear equations for modeled data
l1<-summary(lm(linear_1$theoretical_so2~linear_1$y))
l2<-summary(lm(linear_2$y~linear_1$y))
l3<-summary(lm(linear_3$y~linear_1$y))
l4<-summary(lm(linear_4$y~linear_1$y))
l5<-summary(lm(linear_5$y~linear_1$y))


#Linear model graph 
pl1<-ggplot() +
  geom_line(aes(standard_curve$standardso2, standard_curve$standardso2, color = "Theoretical saturation"))+
  geom_abline(aes(intercept = coef(l1)[1], slope = coef(l1)[2], color= "Patient modeled saturation"), show.legend =TRUE)+
  geom_abline(aes(intercept = coef(l2)[1], slope = coef(l2)[2], color= "Kellman modeled saturation"), show.legend =TRUE)+
  geom_abline(aes(intercept = coef(l3)[1], slope = coef(l3)[2], color= "Severinghaus modeled saturation"), show.legend =TRUE)+
  geom_abline(aes(intercept = coef(l4)[1], slope = coef(l4)[2], color= "Severinghaus (pH adjusted) modeled saturation"), show.legend =TRUE)+
  geom_abline(aes(intercept = coef(l5)[1], slope = coef(l5)[2], color= "Severinghaus (Temp adjusted) modeled saturation"), show.legend =TRUE)+
  labs(x="Oxygen saturation (%)", 
       y="Theoretical saturation (%)",
       title="") +
  scale_color_viridis(name="", discrete=TRUE, option="viridis")+
  theme_bw()

#Linear model graph with all intercepts at 0
pl2<-ggplot() +
  geom_line(aes(standard_curve$standardso2, standard_curve$standardso2, color = "Theoretical saturation"))+
  geom_abline(aes(intercept = 0, slope = coef(l1)[2], color= "Patient modeled saturation"), show.legend =TRUE)+
  geom_abline(aes(intercept = 0, slope = coef(l2)[2], color= "Kellman modeled saturation"), show.legend =TRUE)+
  geom_abline(aes(intercept = 0, slope = coef(l3)[2], color= "Severinghaus modeled saturation"), show.legend =TRUE)+
  geom_abline(aes(intercept = 0, slope = coef(l4)[2], color= "Severinghaus (pH adjusted) modeled saturation"), show.legend =TRUE)+
  geom_abline(aes(intercept = 0, slope = coef(l5)[2], color= "Severinghaus (Temp adjusted) modeled saturation"), show.legend =TRUE)+
  labs(x="Oxygen saturation (%)", 
       y="Theoretical sturation (%)",
       title="") +
  labs(x="Oxygen saturation (%)", 
       y="Theoretical saturation (%)",
       title="") +
  scale_color_viridis(name="", discrete=TRUE, option="viridis")+
  theme_bw()



```

```{r}
labels <- data.frame(x = c(1984, 1942), y = c(150, 150))

ph7<-ggplot()+
  geom_histogram(aes(R1$theoretical), fill="#d7191c", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R2$patient), fill="#fdae61", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R3$kellman), fill="#ffffbf", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R4$severinghaus_unadjusted), fill="#482677", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R5$severinghaus_ph), fill="#abd9e9", color="grey10", alpha=1/10, bins = 200)+
  geom_histogram(aes(R6$severinghaus_temp), fill="black", color="grey10", alpha=1/10, bins = 200)+
  geom_text(data = labels, aes(93.8, 700, label = "Patient", angle = 45), size = 3.5, alpha=5/10)+
  geom_text(data = labels, aes(96.24, 1270, label = "Adairs", angle = 45), size = 3.5, alpha=5/10)+
  geom_text(data = labels, aes(95.5, 1150, label = "Kellman ", angle = 45), size = 3.5, alpha=5/10)+
  geom_text(data = labels, aes(96.56, 1278, label = "Severinghaus", angle = 45), size = 3.5, alpha=5/10)+
  geom_text(data = labels, aes(95.55, 1900, label = "Severinghaus Temp and pH", angle = 45), size = 3.5, alpha=5/10)+
  labs(x="Oxygen saturation (%)", 
       y="Count",
       title="")+ 
  theme_classic()
```


#Format graphs

```{r}

gar1<-ggarrange(p1, pcd1,
          p2, pcd2, 
          labels = c("A", "", "B"),
          ncol = 2, nrow = 2)

gar2<-ggarrange(p3, pcd3,
          p4, pcd4,
          p5, pcd5,
          labels = c("C", "", "D", "", "E"),
          ncol = 2, nrow = 3)

```

#Save plots into as high res tiff

```{r}
#save graphs
tiff("mimic_1.tiff", units="in", width=7.5, height=5.75, res=600)
gar1
dev.off()

tiff("mimic_2.tiff", units="in", width=7.5, height=5.75, res=600)
gar2
dev.off()

tiff("mimic_3.tiff", units="in", width=7.5, height=5.75, res=600)
ph7
dev.off()
```





#Aditional variables
#p50 , po2, spo2 and sofa 

```{r}
#for standard curve
standard_curve %>% select(po2seq, standardso2) %>% summary()
standard_curve %>% select(po2seq, standardso2) %>% filter(standardso2>=50 & standardso2<51) %>% summary()
theoryp50<-df %>% select(PO2, theoretical_so2) %>% filter(theoretical_so2>=50 & theoretical_so2<51)
#for raw data
df %>% select(PO2, SPO2, SOFA) %>% summary()
df %>% select(PO2, SPO2) %>% filter(SPO2>=50 & SPO2<51) %>% summary(ttest=t.test())
rawp50<-df %>% select(PO2, SPO2) %>% filter(SPO2>=50 & SPO2<51)
#statistical test
rawp50<-df %>% select(PO2, SPO2) %>% filter(SPO2>=50 & SPO2<51)
wilcox.test(theoryp50$PO2, rawp50$PO2, alternative = "two.sided")
# SEVERINGHAUS RAW data
df %>% select(PO2, SSO2, SOFA) %>% summary()
df %>% select(PO2, SSO2, SOFA) %>% filter(SSO2>=50 & SSO2<51) %>% summary()
sevp50<-df %>% select(PO2, SSO2) %>% filter(SSO2>=50 & SSO2<51)
wilcox.test(theoryp50$PO2, sevp50$PO2, alternative = "two.sided")
# SEVERINGHAUS ph corrected data
df %>% select(LNSPO2VPH,SS02PH, SOFA) %>% summary()
df %>% select(LNSPO2VPH,SS02PH, SOFA) %>% filter(SS02PH>=50 & SS02PH<51) %>% summary()
php50<-df %>% select(LNSPO2VPH, SS02PH) %>% filter(SS02PH>=50 & SS02PH<51)
#wilcox.test(theoryp50$PO2, php50$LNSPO2VPH, alternative = "two.sided") not enough observations
# SEVERINGHAUS temperature corrected data
df %>% select(lnSPO2V,SSO2T, SOFA) %>% summary()
df %>% select(lnSPO2V,SSO2T, SOFA) %>% filter(SSO2T>=50 & SSO2T<51) %>% summary()
tp50<-df %>% select(lnSPO2V, SSO2T) %>% filter(SSO2T>=50 & SSO2T<51)
wilcox.test(theoryp50$PO2, tp50$SSO2T, alternative = "two.sided")
# kellman
df %>% select(KPO2V, KSO2, SOFA) %>% summary()
df %>% select(KPO2V, KSO2, SOFA) %>% filter(KSO2>=50 & KSO2<51) %>% summary()
kp50<-df %>% select(KPO2V, KSO2) %>% filter(KSO2>=50 & KSO2<51)
wilcox.test(theoryp50$PO2, kp50$KSO2, alternative = "two.sided")
```

#p90

```{r}
#for standard curve
standard_curve %>% select(po2seq, standardso2) %>% summary()
standard_curve %>% select(po2seq, standardso2) %>% filter(standardso2>=90 & standardso2<91) %>% summary()
theoryp90<-df %>% select(PO2, theoretical_so2) %>% filter(theoretical_so2>=90 & theoretical_so2<91)
#for raw data
df %>% select(PO2, SPO2, SOFA) %>% summary()
df %>% select(PO2, SPO2) %>% filter(SPO2>=90 & SPO2<91) %>% summary(ttest=t.test())
rawp90<-df %>% select(PO2, SPO2) %>% filter(SPO2>=90 & SPO2<91)
#statistical test
rawp90<-df %>% select(PO2, SPO2) %>% filter(SPO2>=90 & SPO2<91)
wilcox.test(theoryp90$PO2, rawp90$PO2, alternative = "two.sided")
# SEVERINGHAUS RAW data
df %>% select(PO2, SSO2, SOFA) %>% summary()
df %>% select(PO2, SSO2, SOFA) %>% filter(SSO2>=90 & SSO2<91) %>% summary()
sevp90<-df %>% select(PO2, SSO2) %>% filter(SSO2>=90 & SSO2<91)
wilcox.test(theoryp90$PO2, sevp90$PO2, alternative = "two.sided")
# SEVERINGHAUS ph corrected data
df %>% select(LNSPO2VPH,SS02PH, SOFA) %>% summary()
df %>% select(LNSPO2VPH,SS02PH, SOFA) %>% filter(SS02PH>=90 & SS02PH<91) %>% summary()
php90<-df %>% select(LNSPO2VPH, SS02PH) %>% filter(SS02PH>=90 & SS02PH<91)
wilcox.test(theoryp90$PO2, php90$LNSPO2VPH, alternative = "two.sided") 
# SEVERINGHAUS temperature corrected data
df %>% select(lnSPO2V,SSO2T, SOFA) %>% summary()
df %>% select(lnSPO2V,SSO2T, SOFA) %>% filter(SSO2T>=90 & SSO2T<91) %>% summary()
tp90<-df %>% select(lnSPO2V, SSO2T) %>% filter(SSO2T>=90 & SSO2T<91)
wilcox.test(theoryp90$PO2, tp90$SSO2T, alternative = "two.sided")
# kellman
df %>% select(KPO2V, KSO2, SOFA) %>% summary()
df %>% select(KPO2V, KSO2, SOFA) %>% filter(KSO2>=90 & KSO2<91) %>% summary()
kp90<-df %>% select(KPO2V, KSO2) %>% filter(KSO2>=90 & KSO2<91)
wilcox.test(theoryp90$PO2, kp90$KSO2, alternative = "two.sided")
```
